




The objective of this chapter is to study the problem of calibration and consistency in presence of adversaries. We study, in Section~\ref{sec:calibration}, the problem of calibration in the adversarial setting and provide both necessary and sufficient conditions for a loss to be calibrated in this setting. It also worth noting that our results are easily extendable to  $\mathcal{H}$-calibration (see Section~\ref{xxx}). One on the main takeaway of our analysis is that  no convex surrogate loss can be calibrated in the adversarial setting. We however characterize a set of non-convex loss functions, namely \emph{shifted odd functions} that solve the calibration problem in the adversarial setting. Finally, we focus on the problem of consistency in the adversarial setting in Section~\ref{sec:consistency}. Based on min-max arguments, we provide insights that might help paving a way to prove consistency of shifted odd functions in the adversarial setting. Specifically, we proved strong duality results for these losses and show tight links with the $0/1$-loss. From these insights, we are able to provide a close but weaker property to consistency.

\paragraph{Notations.} Let us consider a classification task with input space $\mathcal{X}$ and output space $\mathcal{Y}=\{-1,+1\}$. Let $(\mathcal{X},d)$ be a proper Polish (i.e. completely separable) metric space representing the inputs space. For all $x\in\mathcal{X}$ and $\delta>0$, we denote $B_\delta(x)$ the closed ball of radius $\delta$ and center $x$. We also assume that for all $x\in\mathcal{X}$ and $\delta>0$,  $B_\delta(x)$ contains at least two points\footnote{For instance, for any norm $\lVert\cdot\rVert$,  $(\mathbb{R}^d,\lVert \cdot \rVert)$ is a Polish metric space satisfying this property.}. Let us also endow $\mathcal{Y}$ with the trivial metric  $d'(y,y') = \mathbf{1}_{y\neq y'}$. Then the space $(\mathcal{X}\times\mathcal{Y},d\oplus d')$ is a proper Polish space. For any Polish space $\mathcal{Z}$, we denote $\mathcal{M}_+^1(\mathcal{Z})$ the Polish space of Borel probability measures on $\mathcal{Z}$. We will denote $\mathcal{F}(\mathcal{Z})$ the space of real valued Borel measurable functions on $\mathcal{Z}$. Finally, we denote $\bar{\RR}:=\RR\cup\{\-\infty,+\infty\}$. Moreover, we take back the definitions introduced in Section~\ref{sec:rw-q2}.



%However, there is no guarantee that the adversarial version of the surrogate loss exhibit the same consistency property toward the adversarial $0/1$ loss. In the standard case,  most results  come from the reduction the global problem of consistency to a local pointwise problem  of \textit{calibration} for each input. This is not possible in the adversarial setting, because the new losses are intrinsically non-pointwise and the optimum of a given input may depend of another one.





